---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

```{r}
library(dplyr)
library(rvest)
library(magrittr)
library(rsalad)
```


# Retrieve population of countries

The last piece of information I want for the analysis is the population size of each country. There are obviously many sources for such information. I could have used Jenny Bryan's Gapminder package, for example, but a better source I found was from GeoNames since they have the 2-letter code for each country, so it'll be easier to merge with our data.

Once again, I will use Hadley Wickham's rvest package, this time to scrape a table of country populations from GeoNames. I'm still excited by how easy this is!

```{r}
countryPopUrl <- "http://www.geonames.org/countries/"
countryPopulationPage <- read_html(countryPopUrl)
countryPop <-
  countryPopulationPage %>%
  html_nodes("table") %>%
  extract2(2) %>%
    html_table

knitr::kable(head(countryPop))
```
```{r}
countryPop %<>% 
    set_colnames(sub("ISO-3166alpha2", "countryCode", colnames(.)) %>%
                 tolowerfirst) %>%
  
    mutate(population = as.numeric((gsub(",", "", population)))) %>%
  select(countryCode, population)
knitr::kable(head(countryPop))
```

# Session info
```{r}
sessionInfo()
```

